Spatial and temporal decay of SARS-CoV-2 environmental contamination of the healthcare environment

### 2021-09-16 manuscript.Rmd
Quick Manuscript Correction

****ADD Discussion of our earlier ICHE paper

We sought to define the impact of time and distance on SARS-CoV-2 surface contamination through longitudinal sampling of  within COVID-19 patient rooms. We found that distance from the patient's bed did not inform the probability of detection of SARS-CoV-2 RNA. We had hypothesized that the probability of RNA detection would decrease with distance. This hypothesis was informed by our prior work with bacterial deposition within patient rooms, where some taxa have decreased probability of detection as distance changes within patient room [@KElly2021]. However, we found that SARS-CoV-2 detection did not vary significantly among any surface group as distance varied. This is most likely related to two factors 1) the relatively small size of patient rooms and 2) the physical spread of viral RNA through touch. We observed large differences in the probability of detecting surface contamination by surface type. Floors had high probability of contamination, which did not vary throughout the room. This is likely related to both the role of patient and healthcare personnel (HCP) foot traffic in addition to the physical action of mopping which may spread body fluids containing viral RNA within the patient's room. This finding matches previous description of high probability of floor contamination relative to other surfaces [@Chia2020]. In contrast to floors, walls had significantly lower probability of contamination. This may be due to the decreased settling of respiratory droplets on vertical surfaces, or the decreased role of touch or secondary contamination through cleaning.

>>>>

We sought to define the impact of time and distance on SARS-CoV-2 surface contamination through longitudinal sampling of  within COVID-19 patient rooms. We found that distance from the patient's bed did not inform the probability of detection of SARS-CoV-2 RNA. We had hypothesized that the probability of RNA detection would decrease with distance. This hypothesis was informed by our prior work with bacterial deposition within patient rooms, where some taxa have decreased probability of detection as distance changes within patient room [@Kelly2021]. However, we found that SARS-CoV-2 detection did not vary significantly among any surface group as distance varied. This is most likely related to two factors 1) the relatively small size of patient rooms and 2) the physical spread of viral RNA through touch. We observed large differences in the probability of detecting surface contamination by surface type. Floors had high probability of contamination, which did not vary throughout the room. This is likely related to both the role of patient and healthcare personnel (HCP) foot traffic in addition to the physical action of mopping which may spread body fluids containing viral RNA within the patient's room. This finding matches previous description of high probability of floor contamination relative to other surfaces [@Chia2020]. In contrast to floors, walls had significantly lower probability of contamination. This may be due to the decreased settling of respiratory droplets on vertical surfaces, or the decreased role of touch or secondary contamination through cleaning.
